---
title: "Assignment 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
```

### Building the book

We got two different versions of the book from two different compilations on two different machines

Version 1:The last word on page 21 is "factor" (Last line: take, even if you don’t see all values in a given dataset. Using a factor)

Version 2: The last word on page 21 is "may" (Last line: Factors are useful when you know the possible values a variable may)

### tmax_clean Function

#### Function for returning cleaned Weather Station data

The compressed weather station [data](https://www1.ncdc.noaa.gov/pub/data/ushcn/v2.5/) containing the tmax temperature files must be unzipped in the same directory as this file. The function takes stationid as argument and returns a data structure with max temperature (column Temp_Max) for each Month (column Month). (Months: 1=January, 2=February and so on)

```{r}
tmax_clean <- function(stationid){

temps <- read_fwf(paste0("ushcn.v2.5.5.20170118/",stationid,".raw.tmax"),
                  col_positions = fwf_positions(
                    start=c( 1,13, rep(16+9*0:11, each=4) + c(1,7,8,9)), 
                    end  =c(11,16, rep(16+9*0:11, each=4) + c(6,7,8,9))))
names(temps) <- c("Station", "Year", 
                  paste0(rep(c("Value","DMflag", "QCflag", "DSflag"), 12), 
                         rep(1:12, each=4)))

station <- temps %>% 
  gather(key="Month", value="Temp_Max", starts_with("Value"))
station$Month <- as.numeric(gsub("Value", "", station$Month))
station$Temp_Max <- replace(station$Temp_Max, 
                            station$Temp_Max == -9999, NA)
return(station)
}
```

#### Calling the Function
Here we are passing the station ID to the tmax_clean function and collecting the cleaned data structure. The Maximum Temperatures are plotted against the Year separately for each month.
```{r message=FALSE}
library(tidyverse)
st<-tmax_clean("USH00300937")
st %>% ggplot(aes(x = Year, Temp_Max/100)) + geom_point() + facet_wrap(~Month, nrow=2)
```

